Load libraries
```{r libraries, echo = TRUE}
library(plyr)
```

Code to Read the data
```{r reading data, echo = TRUE}


activity <- read.csv("activity.csv", stringsAsFactors = FALSE)
```

Code that formats the data into an analyzable table
```{r format data, echo = TRUE}
activity$date <- as.factor(activity$date)

#create a new set that has dates formated into a date object
activity_d <- activity
activity_d$date <- strptime(activity$date, format = "%Y-%m-%d")
```

Total Steps taken per day
```{r total steps, echo = TRUE}
step_sum <- aggregate(activity$steps, by = list(activity$date), sum, na.rm = TRUE)
hist(as.vector(rep(strptime(step_sum$Group.1, format = "%Y-%m-%d"), step_sum$x)), 
     breaks = strptime(step_sum$Group.1, format = "%Y-%m-%d"), 
     freq = TRUE, 
     xlab = "Date", 
     ylab = "Total Number of Steps", 
     main = "Total Steps Taken per Day")
```

Mean and median of total number of steps taken per day
```{r mean and median, echo = TRUE}
activity_rm <- activity[is.na(activity$steps) == FALSE, ]
activity_sums <- aggregate(activity_rm$steps, by = list(activity_rm$date), sum)
summary(activity_sums$x)
```

Time series of Number of steps taken in each interval. Each interval is the average of all days in that interval.
```{r time series, echo = TRUE}
time_series <- aggregate(activity$steps, by = list(activity$interval), mean, na.rm = TRUE)
plot(time_series,
     type = "l",
     xlab = "Time (min)",
     ylab = "Average steps taken")
```

Interval that contains the maximum number of steps
```{r max steps, echo = TRUE}
time_series[which.max(time_series$x), ]
```

Percentage of NA's
```{r percent NAs, echo = TRUE}
sum(is.na(activity$steps))/nrow(activity)
```

Use mean imputation to impute missing values
```{r impute, echo = TRUE}
activity_noNA <- activity[is.na(activity$steps) == FALSE, ]
activity_NA <- activity[is.na(activity$steps) == TRUE, ]
names(time_series) <- c("interval", "x")
imputed <- join(time_series, activity_NA, by = 'interval', type = 'right', match = 'all')
imputed <- data.frame(cbind(imputed$x, as.character(imputed$date), imputed$interval))
names(imputed) <- names(activity_NA)
activity_imputed <- data.frame(rbind(activity_noNA, imputed))
activity_imputed$steps <- as.numeric(activity_imputed$steps)
```

Total Steps taken per day using imputed data
```{r total steps imputed, echo = TRUE}
step_sum2 <- aggregate(activity_imputed$steps, by = list(activity_imputed$date), sum, na.rm = TRUE)
hist(as.vector(rep(strptime(step_sum2$Group.1, format = "%Y-%m-%d"), step_sum2$x)), 
     breaks = strptime(step_sum2$Group.1, format = "%Y-%m-%d"), 
     freq = TRUE, 
     xlab = "Date", 
     ylab = "Total Number of Steps", 
     main = "Total Steps Taken per Day with Mean Imputation")
```

Mean and median of total number of steps taken per day using imputed data
```{r mean and median imputed, echo = TRUE}
activity_sums2 <- aggregate(activity_imputed$steps, by = list(activity_imputed$date), sum)
summary(activity_sums2$x)
```

Make a new column indicating whether the day is a weekday or weekend
```{r day types, echo = TRUE}
day_type <- character()
activity_imputed$date <- strptime(activity_imputed$date, format = "%Y-%m-%d")
date_vector <- activity_imputed$date
nrows <- nrow(activity_imputed)
for (i in 1:17568) {
  if (weekdays(date_vector)[i] %in% c("Saturday", "Sunday")) {
    day_type[i] <- "weekend"
  } else {
    day_type[i] <- "weekday"
  } 
}
activity_imputed$day_type <- day_type
activity_imputed$day_type <- as.factor(activity_imputed$day_type)
```

Panel plot for weekday vs weekends
```{r weekday vs weekends, echo = TRUE}
activity_imputed$steps <- as.numeric(activity_imputed$steps)
activity_imputed$interval <- as.numeric(activity_imputed$interval)
par(mfrow = c(2:1))
weekdays <- subset(activity_imputed, activity_imputed$day_type == "weekday")
weekends <- subset(activity_imputed, activity_imputed$day_type == "weekend")
weekdays_agg <- aggregate(weekdays$steps, by = list(weekdays$interval), mean)
weekends_agg <- aggregate(weekends$steps, by = list(weekends$interval), mean)
plot(weekdays_agg, main = "Weekdays", type = "l", xlab = "Interval", ylab = "Number of steps")
plot(weekends_agg, main = "Weekends", type = "l", xlab = "Interval", ylab = "Number of steps")
```
